Including EW and mixed QCD-EW corrections into PDF fits

### 2020-07-07 paper/sec-pineappl.tex
Add corrections to Marco's corrections

In this paper we present a new library called \textsc{PineAPPL}, which stores phase-space weights of a Monte Carlo (MC) integration independently from the chosen PDFs,
including higher-order corrections.
This separates the task of producing predictions for observables into two steps: 1) the generation of the \emph{grids}, which is how the generated files are called, and 2) the convolution of those grids with a set of PDFs.
The advantage of this method is that the time-consuming step 1) has to be done only once and it allows for very quick convolutions of a given grids with multiple PDF sets.

The convolution is typically done in a few seconds or less, which offers at least two applications:
\begin{enumerate}
\item the study of PDF-dependence of observables; e.g.\ PDF set comparisons, PDF uncertainties for different PDF set, $\alphas$ variations, etc., and
\item the determination of PDF sets themselves; the grids together with the corresponding experimental data constitute two important ingredients for a PDF fit.
\end{enumerate}
This puts \textsc{PineAPPL} in the same category as \textsc{APPLgrid}~\cite{Carli:2010rw} and \textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}. With
respect to the capabilities of these computer codes, \textsc{PineAPPL} makes it possible for the first time to include also higher-order corrections due to electroweak (EW)
(and in general mixed QCD-EW) effects. Documenting this extension, and how to interface \textsc{PineAPPL} with a general-purpose matrix-element generator, is the main interest of this paper.

In particular, the following features are supported:
\begin{itemize}
\item The inclusion of pertrubative corrections (at fixed-order, i.e. without parton-shower matching) with any given set of powers of $\alpha$, $\alphas$, in particular 
    including mixed QCD--EW corrections.
\item The support for non-coloured initial-state partons, such as photon-initiated contributions.
In fact, \textsc{PineAPPL} allows arbitrary initial-state combinations, e.g.\ leptonic initial states~\cite{Bertone:2015lqa,Buonocore:2020nai}.
\item The estimate of theory uncertainty via variations of the renormalisation and factorisation scale (the electroweak coupling is assumed to be scale independent, 
    consistently with the most commonly used renormalisation schemes).
%\item Support for all-order predictions coming from a resummation calculation or a photon-/parton-shower, which are important for some observables (see section~\ref{sec:results}),
\end{itemize}
On a more technical level, we further point out the following differences and improvements in \textsc{PineAPPL}:
\begin{itemize}
    \item \textsc{APPLgrid} typically required a substantial amount of memory (close to \SI{80}{\giga\byte}), if a grid for a 
        differential distributions with more than 10 bins was generated.
The memory usage is substantially reduced (roughly \SI{1}{\giga\byte}) after \enquote{optimisation} of the grids, i.e.\ an optimisation of the grid representation in memory.
However, this required a two-step procedure: first, to produce an unoptimised grid in order to identify those parts where
the cross section is either zero or extremely suppressed; second, after those parts are removed, to fill
an optimised grid with a a small number of bins.
\textsc{PineAPPL} avoids this by using a more space-efficient representation from the start.
\item \textsc{PineAPPL} offers an easy-to-use interface written in the C programming language, to allow MC integrators to read and write \textsc{PineAPPL} grids.
C was chosen because it can be easily interfaced with both Fortran and C++, in which most MC integrators are written.
The interface consists of roughly 30 functions, among which only a handful are needed in practice.
See appendix~\ref{app:example-program} for an example.
%\textsc{PineAPPL} itself is written in Rust (see appendix~\ref{app:installation} for installation instructions).
\item The shell command \texttt{pineappl}, which performs convolutions on the c
    ommand line (see appendix~\ref{app:pineappl-demo}), without requiring the user to write a new program.
In addition to convolutions, it can also print how the luminosity function is constructed, which perturbative orders are stored, their size, the size of each partonic channel, etc., separately for each bin.
%\item Possibility to import \textsc{APPLgrids} and \textsc{fastNLO} tables.
\end{itemize}

We have interfaced \textsc{PineAPPL} with the new version (v3+) of
\textsc{mg5\_aMC@NLO}~\cite{Alwall:2014hca,Frederix:2018nkq}\footnote{It can be downloaded 
    from \url{https://code.launchpad.net/~amcblast/+junk/3.0.2}, and will soon be merged in the official code release}, where it will replace the \textsc{aMCfast}~\cite{Bertone:2014zva} interface.
Its usage is very similar to that of \textsc{mg5\_aMC v2}+\textsc{aMCfast}+\textsc{APPLgrid}.
In practice, using \textsc{PineAPPL} leads to substantially faster run times, in particular for simple processes, because the grids do not need to be optimised 
and their combination is faster.
See appendix~\ref{app:sample-runcard} for an example of a runcard.

>>>>

In this paper we present a new library called \textsc{PineAPPL}, which stores phase-space weights of a Monte Carlo (MC) integration of a fixed-order calculation independently from the chosen PDFs.
This separates the task of producing predictions for observables into two steps: 1) the generation of the \emph{grids}, which is how the generated files are called, and 2) the convolution of those grids with a set of PDFs.
The advantage of this method is that the time-consuming step 1) has to be done only once and, therefore, allows for very quick convolutions of a given grid with multiple PDF sets.

The convolution is typically done in a few seconds or less, which offers at least two applications:
\begin{enumerate}
\item the study of PDF-dependence of observables; e.g.\ PDF set comparisons, PDF uncertainties for different PDF set, $\alphas$ variations, etc., and
\item the determination of PDF sets themselves; the grids together with the corresponding experimental data constitute two important ingredients for a PDF fit.
\end{enumerate}
This puts \textsc{PineAPPL} in the same category as \textsc{APPLgrid}~\cite{Carli:2010rw} and \textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}.
With respect to the capabilities of these computer codes, \textsc{PineAPPL} makes it possible for the first time to include also higher-order corrections due to electroweak (EW), and in general mixed QCD--EW, effects.
Documenting this extension, and how to interface \textsc{PineAPPL} with a general-purpose matrix-element generator, is the main interest of this paper.

In particular, the following features are supported:
\begin{itemize}
\item The inclusion of perturbative corrections (fixed-order, i.e.\ without parton-shower matching) with any given set of powers of $\alpha$, $\alphas$, in particular including mixed QCD--EW corrections.
\item The support for non-coloured initial-state partons, such as photon-initiated contributions.
In fact, \textsc{PineAPPL} allows arbitrary initial-state combinations, e.g.\ leptonic initial states~\cite{Bertone:2015lqa,Buonocore:2020nai}.
\item The estimate of theory uncertainty via variations of the renormalisation and factorisation scale (the electroweak coupling is assumed to be scale-independent, consistently with the commonest renormalisation schemes).
%\item Support for all-order predictions coming from a resummation calculation or a photon-/parton-shower, which are important for some observables (see section~\ref{sec:results}),
\end{itemize}
On a more technical level, we further point out the following differences and improvements in \textsc{PineAPPL}:
\begin{itemize}
\item \textsc{APPLgrid} typically required a substantial amount of memory (close to \SI{120}{\giga\byte} for Drell--Yan), if a grid for a differential distributions with more than 10 bins was generated.
The memory usage is substantially reduced (roughly \SI{1}{\giga\byte}) after \enquote{optimisation} of the grids, i.e.\ an optimisation of the grid representation in memory.
However, this required a two-step procedure: first, to produce an unoptimised grid in order to identify those parts where the cross section is either zero or extremely suppressed; second, after those parts are removed, to fill an optimised grid with a small number of bins.
\textsc{PineAPPL} avoids this by using a more space-efficient representation from the start.
\item \textsc{PineAPPL} offers an easy-to-use interface written in the C programming language, to allow MC integrators to read and write \textsc{PineAPPL} grids.
C was chosen because it can be easily interfaced with both Fortran and C++, in which most MC integrators are written.
The interface consists of roughly 30 functions, among which only a handful are needed in practice.
See appendix~\ref{app:example-program} for an example.
%\textsc{PineAPPL} itself is written in Rust (see appendix~\ref{app:installation} for installation instructions).
\item The shell command \texttt{pineappl}, which performs convolutions on the command line (see appendix~\ref{app:pineappl-demo}), without requiring the user to write a new program.
In addition to convolutions, it can also print how the luminosity function is constructed, which perturbative orders are stored, their size, the size of each partonic channel, etc., separately for each bin.
%\item Possibility to import \textsc{APPLgrids} and \textsc{fastNLO} tables.
\end{itemize}

We have interfaced \textsc{PineAPPL} with the new version (v3+) of \textsc{mg5\_aMC@NLO}~\cite{Alwall:2014hca,Frederix:2018nkq}\footnote{It can be downloaded from \url{https://code.launchpad.net/~amcblast/+junk/3.0.2}, and will soon be merged in the official code release.}, where it will replace the \textsc{aMCfast}~\cite{Bertone:2014zva} interface.
Its usage is very similar to that of \textsc{mg5\_aMC v2}+\textsc{aMCfast}+\textsc{APPLgrid}.
In practice, using \textsc{PineAPPL} leads to substantially faster run times, in particular for simple processes, because the grids do not need to be optimised 
and their combination is faster.
See appendix~\ref{app:sample-runcard} for an example of a runcard.
